---
title: "Gan_Optimization"
output: html_document
date: "2024-07-28"
---

# 2 Initialize library and directory ~~~~~~~~~~~~~~~~~~~~~

```{r, initialize}
require(readxl)
require(xts)
require(readr)
require(tidyverse)
require(broom)
require(stargazer)
require(tidyquant)
require(purrr)
require(tidyr)
require(dplyr)
require(quantmod)
require(PerformanceAnalytics)
require(PortfolioAnalytics)
require(TTR)
require(DEoptim)
require(ROI)
#require(qmao)
require(zoo)
library(reshape2)
library(quantmod)
require(quadprog)
require(corrplot)
require(ggplot2)
require(foreach)
#install.packages("ROI.plugin.quadprog")
#install.packages("ROI.plugin.glpk")
library(ROI.plugin.quadprog)
library(ROI.plugin.glpk)
```

```{r, Start Here}
#path <- "/Users/rgkeat/downloads/tradingProgramming/port1.xlsx"
#port1 <- read_xlsx(path)
#path <- "/Users/rgkeat/downloads/tradingProgramming/port2.xlsx"
#port2 <- read_xlsx(path)
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port1.xlsx"
port1 <- read_xlsx(path)
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port2.xlsx"
port2 <- read_xlsx(path)
```
# 2 Transform ~~~~~~~~~~~~~~~~~~~~~
```{r}
rfRate <- 0.06275 
```


```{r}
port2 <- merge(port1, port2, by = "date", all = TRUE)
```

```{r}
# Convert into df
port1 <- data.frame(port1)
port2 <- data.frame(port2)
head(port1)

```


```{r}
# Convert to numeric except date
port1[, -1] <- lapply(port1[, -1], function(x) as.numeric(as.character(x)))
port2[, -1] <- lapply(port2[, -1], function(x) as.numeric(as.character(x)))
```

```{r}
# Drop missing
port1 <- na.omit(port1)
port2 <- na.omit(port2)
#port1 <- port1[complete.cases(port1), ]
#port2 <- port2[complete.cases(port2), ]
sum(is.na(port1))
sum(is.na(port2))
```




```{r}
port1 <- xts(port1[,-1], order.by = as.Date(port1$date))
port2 <- xts(port2[,-1], order.by = as.Date(port2$date))
head(port1)
```

```{r}
port1ret <- PerformanceAnalytics::Return.calculate(port1,method="log")
port1ret <- port1ret[-1]
```

```{r}
port2ret <- PerformanceAnalytics::Return.calculate(port2,method="log")
port2ret <- port2ret[-1]
```

```{r}
sum(is.na(port1ret))
sum(is.na(port2ret))
```


# 3 Exploratory Data Analysis (EDA) for both Portfolios ~~~~~~~~~~~~~~~~~~~~~
## 3.1 EDA Port1 and Port2

## 3.1.1 Descriptive Statistics - mean, std dev, skewness and kurtosis of asset returns
```{r,port1 Table}
riskTable10 <- data.frame(port1ret)
riskTable10 <- riskTable10[complete.cases(riskTable10), ]
rfRate <- 0.06275  
means10 <- colMeans(riskTable10)
risks10 <- apply(riskTable10, 2, sd)
sharpe10 <- ifelse(risks10 != 0, (means10 - rfRate) / risks10, NA)
riskReturnBasket10 <- data.frame(Asset = names(means10),Mean_Return = means10,Risk = risks10,Sharpe_Ratio = sharpe10)
print(riskReturnBasket10)
```

```{r,port2 Table}
riskTable20 <- data.frame(port2ret)
riskTable20 <- riskTable20[complete.cases(riskTable20), ]
means20 <- colMeans(riskTable20)
risks20 <- apply(riskTable20, 2, sd)
sharpe20 <- ifelse(risks20 != 0, (means20 - rfRate) / risks20, NA)
riskReturnBasket20 <- data.frame(Asset = names(means20),Mean_Return = means20,Risk = risks20,Sharpe_Ratio = sharpe20)
print(riskReturnBasket20)
```

```{r}
summary_returns_10 <- table.Stats(port1ret)
print(summary_returns_10)
```
```{r}
if(any(port2ret <= 0)) {
    print("Data contains non-positive values.")
}
```

```{r}
summary_returns_20 <- table.Stats(port2ret)
print(summary_returns_20)
```

## 3.1.2 Correlation matrix of asset returns.

```{r}
correlation_matrix_returns_10 <- cor(port1ret, use = "complete.obs")
print(correlation_matrix_returns_10)
```
```{r}
correlation_matrix_returns_20 <- cor(port2ret, use = "complete.obs")
print(correlation_matrix_returns_20)
```

## 3.1.3 Visualization - time series plots, histograms, density plots of asset returns; HEATMAP of the correlation matrix
```{r}
corrplot(correlation_matrix_returns_10, method = "color")
melted_correlation_matrix <- melt(correlation_matrix_returns_10)
ggplot(melted_correlation_matrix, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme_minimal() +
  labs(title = "Correlation Matrix Heatmap", x = "Assets", y = "Assets")
```

```{r}
all_returns <- merge(port1ret, port2ret)
all_returns_corr <- data.frame(port1ret = rnorm(100), port2ret = rnorm(100))
```

```{r}
correlation_matrix <- cor(all_returns_corr)
print(correlation_matrix)
```


```{r}
corrplot(correlation_matrix, method = "color")
melted_correlation_matrix <- melt(correlation_matrix)

ggplot(melted_correlation_matrix, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme_minimal() +
  labs(title = "Correlation Matrix Heatmap", x = "Assets", y = "Assets")
```

```{r}
sum(is.na(port1ret))
sum(is.na(port2ret))
```
```{r}
chart.CumReturns(port1ret, main = "Ret ALL for R/R Port1"
                 ,wealth.index = TRUE
                 ,legend.loc = TRUE
                 ,geometric = FALSE
                 )
```

## Plot Cumulative Returns port1
```{r}
chart.CumReturns(port1ret, main = "Cumulative Ret for R/R Port1"
                 ,wealth.index = TRUE
                 ,legend.loc = TRUE
                 )
```




```{r}
summary(port1ret)
boxplot(port1ret,main="Boxplot",las=2)
```
```{r}
str(port2ret)
```

```{r}
#port1 <- na.omit(port1)
#port2 <- na.omit(port2)
#port1 <- port1[complete.cases(port1), ]
#port2 <- port2[complete.cases(port2), ]
sum(is.na(port1ret))
sum(is.na(port2ret))
```


```{r}
charts.PerformanceSummary(port1ret
                          ,Rf=rfRate
                          ,main = "ALL Performance Port 1"
                          ,geometric=FALSE
#                          ,methods="StdDev"
                          ,plot.engine="ggplot2"
                          )
```
```{r}
charts.PerformanceSummary(port1ret
                          ,Rf=rfRate
                          ,main = "Cumulative Performance Port 1"
#                          ,geometric=FALSE
#                          ,methods="StdDev"
                          ,plot.engine="ggplot2"
                          )
```

# Plot Cumulative Returns port2
```{r}
chart.CumReturns(port2ret
                 ,Rf=rfRate
                 ,main = "ALL Ret for R/R Port2"
                 ,wealth.index = TRUE
                 ,legend.loc = TRUE
                 ,geometric=FALSE
                 ,methods="StdDev"
#                 ,plot.engine="ggplot2"
                 )
```
```{r}
chart.CumReturns(port2ret
                 ,Rf=rfRate
                 ,main = "Cumulative Ret for R/R Port2"
                 ,wealth.index = TRUE
                 ,legend.loc = TRUE
#                 ,geometric=FALSE
                 ,methods="StdDev"
#                 ,plot.engine="ggplot2"
                 )
```

```{r}
charts.PerformanceSummary(port2ret,Rf=rfRate,main="cumulative performance port2")
```
```{r}
charts.PerformanceSummary(port2ret,Rf=rfRate,geometric=FALSE,main="ALL performance port2")
```

# 4. Modeling ~~~~~~~~~~~~~~~~~~~~~

## 4.1 Portfolio 1
### 4.1.1 Optimized Portfolio with No Rebalancing
#### Initiate the portfolio object for fitting the objective, constraints, and other hyperparameters 
```{r}
portf_10 <- portfolio.spec(assets = colnames((port1ret)))
portf_20 <- portfolio.spec(assets = colnames((port2ret)))
```

#### Add constraints port1
```{r}
portf_10 <- add.constraint(portf_10, type = "full_investment")
portf_10 <- add.constraint(portf_10, type = "long_only")
portf_10 <- add.constraint(portf_10, type = "box", min = 0.0, max = 1.01)
```

#### Add constraints port2
```{r}
portf_20 <- add.constraint(portf_20, type = "full_investment")
portf_20 <- add.constraint(portf_20, type = "long_only")
portf_20 <- add.constraint(portf_20, type = "box", min = 0.0, max = 1.01)
```

#### initialize return and risk parameters for port1 specifications
```{r}
# Objective to maximize return
portf_10 <- add.objective(portf_10, type = "return", name = "mean")

# Objective to minimize risk (standard deviation)
portf_10 <- add.objective(portf_10, type = "risk", name = "StdDev")
```

#### initialize return and risk parameters for port2 specifications
```{r}
# Objective to maximize return
portf_20 <- add.objective(portf_20, type = "return", name = "mean")

# Objective to minimize risk (standard deviation)
portf_20 <- add.objective(portf_20, type = "risk", name = "StdDev")
```

#### 4.1.1 Optimized Portfolio with No Rebalancing
```{r}
opt_10 <- PortfolioAnalytics::optimize.portfolio(port1ret
                                                 ,portfolio = portf_10
                                                 ,optimize_method = "ROI"
                                                 ,trace = TRUE
                                                 )
print(opt_10)
```

### 4.2 Optimized Portfolio with Rebalancing port1
```{r}
opt_10_rebalanced <- optimize.portfolio.rebalancing(
  R = port1ret
  ,portfolio = portf_10
  ,optimize_method = "ROI"
  ,rebalance_on = "quarter"
  ,trace = TRUE
)
```

```{r,eval=TRUE}
# Extract and normalize weights
weights_10_rebalanced <- na.omit(extractWeights(opt_10_rebalanced))
```


```{r,eval=FALSE}
# Normalize weights if necessary
#(weights_10_rebalanced <- weights_10_rebalanced / rowSums(weights_10_rebalanced, na.rm = TRUE))
# Extract the objective measures for the single period risk optimization with yearly rebalancing
#(weights_10_rebalanced_objectivemeasures <- extractObjectiveMeasures(opt_10_rebalanced))
```

# Optimize Portfolio - 20 stocks
```{r}
opt_20 <- optimize.portfolio(port2ret, 
                             portfolio = portf_20
                             ,optimize_method = "ROI"
                             ,trace = TRUE
                             )
print(opt_20)
```

# Extract Portfolio Weights for 20 stocks
```{r}
weights_20 <- extractWeights(opt_20)
print(weights_20)
```

# Optimization with Rebalancing - 20 stocks
```{r}
opt_20_rebalanced <- optimize.portfolio.rebalancing(
  R = port2ret
  ,portfolio = portf_20
  ,optimize_method = "ROI"
  ,trace = TRUE
  ,search_size = 1000
  ,rebalance_on = "quarters"
  )
```

```{r}
## Extract the objective measures for the single period risk optimization with Quarterly rebalancing

weights_20_rebalanced <- extractWeights(opt_20_rebalanced)
weights_20_rebalanced_objectivemeasures <- extractObjectiveMeasures(opt_20_rebalanced)
```


## plot optimal weights
```{r}
chart.Weights(opt_10, plot.type = "bar")
chart.Weights(opt_20, plot.type = "bar")
chart.Weights(opt_20_rebalanced, plot.type = "bar")
chart.Weights(opt_10_rebalanced, plot.type = "bar")
```

## port1 chart
```{r}
chart.Weights(opt_10
              ,neighbors = NULL
              ,main = "Optimal Portfolio Weights - 10"
              ,las = 1
              ,xlab = "Stock"
              ,cex.lab = 1
              ,element.color = "darkgray"
              ,cex.axis = 0.8
              ,colorset = c("#e5F5F9", "#99D8C9")
              ,legend.loc = "topright"
              ,cex.legend = 0.5
              ,plot.type = "bar"
              )
```

## port2 chart
```{r}
chart.Weights(opt_20
              ,neighbors = NULL
              ,main = "Optimal Portfolio Weights - 20"
              ,las = 1
              ,xlab = "Stock"
              ,cex.lab = 1
              ,element.color = "darkgray"
              ,cex.axis = 0.8
              ,colorset = c("#e5F5F9", "#99D8C9")
              ,legend.loc = "topright"
              ,cex.legend = 0.5, plot.type = "bar"
              )
```

```{r}
weights_10 <- extractWeights(opt_10)
weights_20 <- extractWeights(opt_20)
```

## Calculate Portfolio Returns (for 10 & 20 stocks and with rebalancing)
```{r}
portfolio_returns_10 <- Return.portfolio(R = port1ret, weights = weights_10)
portfolio_returns_10_rebalanced <- Return.portfolio(R = port1ret, weights = weights_10_rebalanced)

portfolio_returns_20 <- Return.portfolio(R = port2ret, weights = weights_20)
portfolio_returns_20_rebalanced <- Return.portfolio(R = port2ret, weights = weights_20_rebalanced)
```

## Comparison 10 stocks with Rebalancing vs No Rebalancing
```{r}
allopt_10 <- merge(portfolio_returns_10, portfolio_returns_10_rebalanced)
colnames(allopt_10) <- c("Portfolio Returns (No Rebalancing)", "Portfolio Returns (With Rebalancing)")
print(colnames(allopt_10))
```

## Comparison 20 stocks with Rebalancing vs No Rebalancing
```{r}
allopt_20 <- merge(portfolio_returns_20, portfolio_returns_20_rebalanced)
colnames(allopt_20) <- c("Portfolio Returns (No Rebalancing)", "Portfolio Returns (With Rebalancing)")
print(colnames(allopt_20))
```
```{r}
allopt_20
```

## Descriptive Statistics
```{r}
summary_portfolio_returns_10 <- table.Stats(portfolio_returns_10$portfolio.returns)
summary_portfolio_returns_10_rebalanced <- table.Stats(portfolio_returns_10_rebalanced$portfolio.returns)
summary_portfolio_returns_20 <- table.Stats(portfolio_returns_20$portfolio.returns)
summary_portfolio_returns_20_rebalanced <- table.Stats(portfolio_returns_20_rebalanced$portfolio.returns)
```

```{r}
all_port_returns <- merge(portfolio_returns_10, portfolio_returns_10_rebalanced, portfolio_returns_20, portfolio_returns_20_rebalanced)
all_port_returns_corr <- data.frame(
  portfolio_returns_10 = rnorm(100)
  ,portfolio_returns_10_rebalanced = rnorm(100)
  ,portfolio_returns_20 = rnorm(100)
  ,portfolio_returns_20_rebalanced = rnorm(100)
  )
```

```{r}
correlation_matrix <- cor(all_port_returns_corr)
print(correlation_matrix)
```


```{r}
corrplot(correlation_matrix, method = "color")
melted_correlation_matrix <- melt(correlation_matrix)

ggplot(melted_correlation_matrix, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme_minimal() +
  labs(title = "Correlation Matrix Heatmap", x = "Assets", y = "Assets")
```

## Plot Cumulative Returns for 10 stocks
```{r}
chart.CumReturns(portfolio_returns_10, main = "Cumulative Returns - R/R Port10", wealth.index = TRUE)
```

```{r}
charts.PerformanceSummary(portfolio_returns_10)
```

## Plot Cumulative Returns for 20 stocks
```{r}
chart.CumReturns(portfolio_returns_20, main = "Cumulative Returns - R/R Port20", wealth.index = TRUE)
```

```{r}
charts.PerformanceSummary(portfolio_returns_20)
```
## Plot Cumulative Returns for 10 stocks with rebalancing
```{r}
chart.CumReturns(portfolio_returns_10_rebalanced, main = "Cumulative Returns - R/R Port10 w rebalancing", wealth.index = TRUE)
```

```{r}
charts.PerformanceSummary(portfolio_returns_10_rebalanced)
```

## Plot Cumulative Returns for 20 stocks with rebalancing
```{r}
chart.CumReturns(portfolio_returns_20_rebalanced, main = "Cumulative Returns - R/R Port20 w rebalancing", wealth.index = TRUE)
```

```{r}
charts.PerformanceSummary(portfolio_returns_20_rebalanced)
```


## Optimization, Rebalancing, Effient Frontier
## Efficient Frontier for 10 stocks

## Function to generate multiple points on the efficient frontier
```{r}
generate_efficient_frontier_10 <- function(port1ret, portf_10, num_points = 50) {
  eff_frontier <- data.frame(Return = numeric(), Risk = numeric(), SharpeRatio = numeric())
  min_return <- min(colMeans(port1ret, na.rm = TRUE))
  max_return <- max(colMeans(port1ret, na.rm = TRUE))
  
  target_returns <- seq(min_return, max_return, length.out = num_points)
  
  for (target_return in target_returns) {
    tmp_portfolio <- portf_10
    tmp_portfolio <- add.constraint(tmp_portfolio, type = "return", return_target = target_return)
    tmp_portfolio <- add.objective(tmp_portfolio, type = "risk", name = "StdDev")
    opt <- optimize.portfolio (R = port1ret, portfolio = tmp_portfolio, optimize_method = "ROI", trace = TRUE)
    ret <- mean(Return.portfolio(R = port1ret, weights = extractWeights(opt)))
    risk <- StdDev(Return.portfolio(R = port1ret, weights = extractWeights(opt)))
    sharpe <- ret/risk
    eff_frontier <- rbind(eff_frontier, data.frame(Return = ret, Risk = risk, SharpeRatio = sharpe))
  }
  return(eff_frontier)
}
```

## Generate efficient frontier
```{r}
efficient_frontier <- generate_efficient_frontier_10(port1ret, portf_10, num_points = 100)
```

## Plot efficient frontier
```{r}
plot(efficient_frontier$Risk
     ,efficient_frontier$Return
     ,type = "l"
     ,col="black"
     ,lwd = 1
     ,xlab = "Risk (StdDev)"
     ,ylab = "Return"
     ,main = "Efficient Frontier"
     )
points(efficient_frontier$Risk
       ,efficient_frontier$Return
       ,col = "grey"
#       ,pch = 16
)
```


# Efficient Frontier ~~~~~~~~~~~~~~~~~~~~~

```{r}
generate_efficient_frontier_20 <<- function(port2ret, portf_20, num_points = 50) {
  eff_frontier <<- data.frame(Return = numeric(), Risk = numeric(), SharpeRatio = numeric())
  min_return <<- min(colMeans(port2ret, na.rm = TRUE))
  max_return <<- max(colMeans(port2ret, na.rm = TRUE))
  target_returns <- seq(min_return, max_return, length.out = num_points)
  
  for (target_return in target_returns) {
    tmp_portfolio <<- portf_20
    tmp_portfolio <<- add.constraint(tmp_portfolio, type = "return", return_target = target_return)
    tmp_portfolio <<- add.objective(tmp_portfolio, type = "risk", name = "StdDev")
    
    opt <<- optimize.portfolio(R = port2ret, portfolio = tmp_portfolio, optimize_method = "ROI", trace = TRUE)
    ret <<- mean(Return.portfolio(R = port2ret, weights = extractWeights(opt)))
    risk <<- StdDev(Return.portfolio(R = port2ret, weights = extractWeights(opt)))
    
    sharpe <<- ret/risk
    eff_frontier <<- rbind(eff_frontier, data.frame(Return = ret, Risk = risk, SharpeRatio = sharpe))
  }
  
  return(eff_frontier)
}
```

# Generate efficient frontier
```{r}
efficient_frontier <- generate_efficient_frontier_20(port2ret, portf_20, num_points = 100)
```

# Plot efficient frontier
```{r}
plot(efficient_frontier$Risk
     ,efficient_frontier$Return
     ,type = "l"
     ,col="black"
     ,lwd = 2
     ,xlab = "Risk (StdDev)"
     ,ylab = "Return"
     ,main = "Efficient Frontier"
     )
points(efficient_frontier$Risk
       ,efficient_frontier$Return
       ,col = "gray"
#       ,pch = 16
       )
```





```{r}
break
```

# Optimization ~~~~~~~~~~~~~~~~~~~~~

```{r}
port_spec1 <- portfolio.spec(assets = colnames(port1))
port_spec1
View(port_spec1)
```

```{r}
#min=0.99 max1.01 to relax constraint

port_spec1 <- add.constraint(portfolio=port_spec1,type="weight_sum"
                             ,min_sum=0.0
                             ,max_sum=1.0)

```

```{r}
port_spec1 <- add.constraint(portfolio = port_spec1
                             ,type = "long_only")
```

```{r}
port_spec1 <- add.constraint(portfolio = port_spec1
                             ,type = "box"
                             , min=0.0
                             , max=1.0)
```

```{r}
# Can be minimize risk or max return
# For max specify type = "return" or type = "risk"
portMinVar1 <- port_spec1
portMinVar1 <- add.objective(portfolio = portMinVar1
                             ,type= "risk"
                             ,name = "StdDev")
View(portMinVar1)
```

```{r}
minVarOpt1 <- optimize.portfolio(R = port1
                                 , portfolio = portMinVar1
                                 , optimize_method = "random"
                                 )
```

```{r}
(minVarOpt1_Weights <- extractWeights(minVarOpt1))
```

```{r}
minVarOpt1_Return <- Return.portfolio(R = port1, 
                                weights = minVarOpt1_Weights)
```

```{r}
# Quadtratic Programming
quadProg1 <- optimize.portfolio(R = port1
                                 , portfolio = portMinVar1
                                 , optimize_method = "quadprog"
                                 ,verbose=TRUE
                                 )
```

```{r}
(quadProg1_Weights <- extractWeights(minVarOpt1))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Appendix ~~~~~~~~~~~~~~~~~~~~~

```{r}
break
```

## Appendix 1: Code used to transform 20 xlsx Files from Refinitiv Eikon
```{r,eval=FALSE}
# Define portfolios
portfolio1 <- c("acen", "allhc", "apx", "cnpf", "bpi", "cnvrg", "fgen", "jgs", "meg", "pxp")
portfolio2 <- c("agi", "ali", "bdo", "ict", "mac", "wlcon", "nikl", "smc", "cbc", "psei")
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/"
```

```{r,transform1,eval=FALSE}
# 20 downloaded xlsx files acquired from refinitiv eikon
# for loop logic: (1) file name aliased as 'ticker' name (2) drop all fields except exchange date and close and append Ticker name (3) 'ticker' is then mapped to portfolio1 or portfolio2, see list above (4) loop repeats until all tickers are bucketed (5) output is a large list of two portfolios with 10 stocks each
file_list <- list.files(path = path, pattern = "*.xlsx", full.names = TRUE)

df_list_portfolio1 <- list()
df_list_portfolio2 <- list()

for (file in file_list) {
  df <- read_excel(file)
  ticker <- tools::file_path_sans_ext(basename(file)) 
  df <- df %>% select(`Exchange Date`, Close) %>% mutate(Ticker = ticker) 
  
  if (ticker %in% portfolio1) {
    df_list_portfolio1 <- append(df_list_portfolio1, list(df))
  } else if (ticker %in% portfolio2) {
    df_list_portfolio2 <- append(df_list_portfolio2, list(df))
  }
}
df_list_portfolio1
df_list_portfolio2
```

```{r,transform2,eval=FALSE}
# merge the two large list into a two dataframes
port1 <- data.frame(bind_rows(df_list_portfolio1))
port2 <- data.frame(bind_rows(df_list_portfolio2))
port1
port2
```

```{r,transform3, eval=FALSE}
# merge again two form 11 fields with close as the price and ticker as the column
port1 <- port1 %>%
  select(Exchange.Date, Ticker, Close) %>%
  pivot_wider(names_from = Ticker, values_from = Close)
port2 <- port2 %>%
  select(Exchange.Date, Ticker, Close) %>%
  pivot_wider(names_from = Ticker, values_from = Close)

# rename
colnames(port1)[1] <- "date"
colnames(port2)[1] <- "date"

# write to output file for data submission, this file is used below. it is saved in your setwd or working directory
#write.csv(port1, "port1.csv",row.names = FALSE) 
#write.csv(port2, "port2.csv",row.names = FALSE) 


port2
port1
```



## Appendix 2: Other code chunks for computing Returns

```{r,log return,eval=FALSE}
# Log returns w na.omit

logReturns1na <- na.omit(CalculateReturns(port1, method = "log"))
logReturns2na <- na.omit(CalculateReturns(port2, method = "log"))
view(logReturns2na)
```

```{r, log return v2 - Only for single returns,eval=FALSE}
# Log returns for single stocks
logreturns_single1 <- periodReturn(port1,
             period="daily"
             ,subset = NULL
             ,type = "log"
             ,leading = TRUE)
logreturns_single2 <- periodReturn(port2,
             period="daily"
             ,subset = NULL
             ,type = "log"
             ,leading = TRUE)
view(logreturns_single1)
```

```{r,eval=FALSE}
# ROC returns
returns <- ROC(port1, type = "continuous")[-1]
```

```{r,port1 for 2 stocks w 50/50 ratio,eval=FALSE}
logreturns_double1 <- na.omit(port1yr)
logreturns_double1 <- logreturns_double1[,1:2]
```

```{r,eval=FALSE}
logreturns_double2 <- na.omit(port2yr)
logreturns_double2 <- logreturns_double2[,1:2]
```

```{r,port1 for 2 stocks w 50/50 ratio,eval=FALSE}
port1ret <- PerformanceAnalytics::Return.portfolio(logreturns_double1,c(0.5,0.5))
view(port1ret)
colnames(port1ret) <- "port1_50_Ret"
mean(port1ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```

```{r,port2 for 2 stocks w 50/50 ratio,eval=FALSE}

port2ret <- PerformanceAnalytics::Return.portfolio(logreturns_double2,c(0.5,0.5))
view(port2ret)
colnames(port2ret) <- "port2_50_Ret"
mean(port2ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```


## Appendix 3: Handling Infinity and NA

```{r, Inf Checker function,eval=FALSE}
# Code is breaking due to instances of infinity when a 0 is returned and is subsequently divided by a function

# initialize infinity check function
infChecker <- function(data) {
  infIndices <- which(is.infinite(data), arr.ind = TRUE)
  return(infIndices)
}

# var list
dataList <<- list(logreturns_single1
           ,logreturns_single2
           ,logReturns1
           ,logReturns2
           ,logReturns1na
           ,logReturns2na
           ,port1
           ,port2
           ,logreturns_double1
           ,logreturns_double2
           ,port1ret
           ,port2ret
           ,port1yr
           ,port2yr
           ) 
vrList <<- c("logreturns_single1"
           ,"logreturns_single2"
           ,"logReturns1"
           ,"logReturns2"
           ,"logReturns1na"
           ,"logReturns2na"
           ,"port1"
           ,"port2"
           ,"logreturns_double1"
           ,"logreturns_double2"
           ,"port1ret"
           ,"port2ret"
           ,"port1yr"
           ,"port2yr"
           ) 

# Apply f(x)
infIndices <- lapply(dataList, infChecker)
names(infIndices) <- vrList
#print(infIndices)
```

```{r, handling Infinity,eval=FALSE}
# Here from the above code we can see both port1ret and port2ret variable is causing the code to break
# Inf value is breaking code due to division by 0
#port2yrInf <- which(is.infinite(port2ret), arr.ind = TRUE)

lapply(port2ret, infChecker)
lapply(port1ret, infChecker)

```

```{r,eval=FALSE}
port1ret[is.infinite(port1ret)] <- NA
port1ret <- port1ret[complete.cases(port1ret), ]
lapply(port1ret, infChecker)
```

```{r, handling Infinity,eval=FALSE}
# Use NA and Drop
port2ret[is.infinite(port2ret)] <- NA
port2ret <- port2ret[complete.cases(port2ret), ]
lapply(port2ret, infChecker)
```


## Appendix 4: Rebalancing

```{r,port1 for 2 stocks w 50/50 ratio,eval=FALSE}
#logreturns_double1 <- na.omit(port1yr)
#logreturns_double1 <- logreturns_double1[,1:2]
#port1ret <- PerformanceAnalytics::Return.portfolio(logreturns_double1,c(0.5,0.5))
view(port1ret)
colnames(port1ret) <- "port1_50_Ret"
mean(port1ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```

```{r,port2 for 2 stocks w 50/50 ratio,eval=FALSE}
# Re initialize without the Inf values
#logreturns_double2 <- na.omit(port2yr)
#logreturns_double2 <- logreturns_double2[,1:2]
#port2ret <- PerformanceAnalytics::Return.portfolio(logreturns_double2,c(0.5,0.5))
view(port2ret)
colnames(port2ret) <- "port2_50_Ret"
mean(port2ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```

```{r,rebal_port1_doublestocks,eval=FALSE}
port1_50 <- na.omit(port1yr)
port1_50 <- port1_50[,1:2]
port1RetRebal <- Return.portfolio(port1_50, c(0.5,0.5),rebalance_on = "months")
view(port1RetRebal)
colnames(port1RetRebal) <- "port50Rebal_Ret"
```

```{r,eval=FALSE}
head(port1ret,30)
head(port1RetRebal,30)
```

```{r,rebal_port2_doublestocks,eval=FALSE}
port2_50 <- na.omit(port2yr)
port2_50 <- port1_50[,1:2]
port2RetRebal <- Return.portfolio(port2_50, c(0.5,0.5),rebalance_on = "months")
view(port2RetRebal)
colnames(port2RetRebal) <- "port50Rebal_Ret"
```

```{r,eval=FALSE}
head(port2ret,30)
head(port2RetRebal,30)
```

```{r,eval=FALSE}
allport1ret <- cbind(port1ret,port1RetRebal)
colnames(allport1ret) <- c("Non-Rebalanced","Monthly Rebalanced")
table.AnnualizedReturns(allport1ret,Rf=0.1/2016)
```


## Appendix 5: Port specifications code chunks
```{r,eval=FALSE}
# Create initial portfolio specification
portfolio_spec <- portfolio.spec(assets = port1)

# Add constraint for full investment
portfolio_spec <- add.constraint(portfolio = portfolio_spec, 
                                 type = "full_investment")

# Add constraint for long only positions
portfolio_spec <- add.constraint(portfolio = portfolio_spec, 
                                 type = "long_only")
# Add weight constraints (e.g., between 5% and 40% for each asset)
portfolio_spec <- add.constraint(portfolio = portfolio_spec, type = "box", min = 0.0, max = 1.0)

#Use box in your output as per recommendation
#Long only minimum 0 and .05
# 0 is unnecessary if there is 0.5
```




## Appendix 6: Efficient Frontier Models

```{r,eval=FALSE}
#Generate Efficient Frontier



# Function to generate multiple points on the efficient frontier
generate_efficient_frontier <- function(returns, portfolio_spec, num_points = 50) {
  eff_frontier <- data.frame(Return = numeric(), Risk = numeric(), SharpeRatio = numeric())
  
  min_return <- min(colMeans(returns))
  max_return <- max(colMeans(returns))
  
  target_returns <- seq(min_return, max_return, length.out = num_points)
  
  for (target_return in target_returns) {
    tmp_portfolio <- portfolio_spec
    
    # Add the return objective dynamically for each target return
    tmp_portfolio <- add.constraint(tmp_portfolio, type = "return", return_target = target_return)
    tmp_portfolio <- add.objective(tmp_portfolio, type = "risk", name = "StdDev")
    
    opt <- optimize.portfolio(R = returns, portfolio = tmp_portfolio, optimize_method = "ROI", trace = TRUE)
    
    ret <- mean(Return.portfolio(R = returns, weights = extractWeights(opt)))
    risk <- StdDev(Return.portfolio(R = returns, weights = extractWeights(opt)))
    sharpe <- ret / risk
    
    eff_frontier <- rbind(eff_frontier, data.frame(Return = ret, Risk = risk, SharpeRatio = sharpe))
  }
  
  return(eff_frontier)
}

```

```{r,eval=FALSE}
# Generate efficient frontier
efficient_frontier <- generate_efficient_frontier(returns, portfolio_spec, num_points = 100)

# Plot efficient frontier
plot(efficient_frontier$Risk, efficient_frontier$Return, type = "l", col = "blue", lwd = 2,
     xlab = "Risk (StdDev)", ylab = "Return", main = "Efficient Frontier")
points(efficient_frontier$Risk, efficient_frontier$Return, col = "blue", pch = 16)

```


```{r,eval=FALSE}
# Function to extract and calculate returns
calculate_returns <- function(data) {
  if (is.null(data) || nrow(data) == 0 || 
      !"Exchange Date" %in% names(data) || !"Close" %in% names(data)) {
    warning("Invalid data structure")
    return(NULL)
  }
  
  data %>%
    select(`Exchange Date`, Close) %>%
    arrange(`Exchange Date`) %>%
    mutate(Returns = (Close - lag(Close)) / lag(Close)) %>%
    select(`Exchange Date`, Returns) %>%
    na.omit()
}

# Process both portfolios
process_portfolio <- function(portfolio_stocks) {
  portfolio_data <- map(stock_data[portfolio_stocks], calculate_returns)
  portfolio_data <- compact(portfolio_data)  # Remove any NULL results
  
  # Combine all stocks' returns into a single dataframe
  reduce(portfolio_data, function(acc, x) {
    full_join(acc, x, by = "date") %>%
      rename(!!sym(names(which(portfolio_stocks == cur_data()$name))) := Returns)
  })
}

# Calculate returns for both portfolios
portfolio1_combined <- process_portfolio(portfolio1)
portfolio2_combined <- process_portfolio(portfolio2)

# Check the results
print(dim(portfolio1_combined))
print(names(portfolio1_combined))
print(dim(portfolio2_combined))
print(names(portfolio2_combined))
```


```{r,eval=FALSE}
library(dplyr)
library(purrr)
library(tidyr)

# Function to extract and calculate returns
calculate_returns <- function(data) {
  if (is.null(data) || nrow(data) == 0 ||
      !"date" %in% names(data) || !"Close" %in% names(data)) {
    warning("Invalid data structure")
    return(NULL)
  }
  
  data %>%
    select(`Exchange Date`, Close) %>%
    arrange(date) %>%
    mutate(Returns = (Close - lag(Close)) / lag(Close)) %>%
    select(date, Returns) %>%
    na.omit()
}

# Process both portfolios
process_portfolio <- function(portfolio_stocks, stock_data) {
  portfolio_data <- map(stock_data[portfolio_stocks], calculate_returns)
  portfolio_data <- compact(port1ret) # Remove any NULL results
  
  # Combine all stocks' returns into a single dataframe
  combined_data <- reduce(portfolio_data, function(acc, x) {
    full_join(acc, x, by = "Exchange Date")
  })
  
  combined_data
}

stock_data <- list(
  stock1 = port1ret(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 100, 200)),
  stock2 = port1ret(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 50, 150)),
  stock3 = port1ret(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 200, 300))
)

portfolio1 <- c("stock1", "stock2")
portfolio2 <- c("stock2", "stock3")

# Calculate returns for both portfolios
portfolio1_combined <- process_portfolio(portfolio1, stock_data)
portfolio2_combined <- process_portfolio(portfolio2, stock_data)

# Check the results
print(dim(portfolio1_combined))
print(names(portfolio1_combined))
print(dim(portfolio2_combined))
print(names(portfolio2_combined))

```


```{r,eval=FALSE}
# Step 5: Generate the efficient frontier
generate_efficient_frontier <- function(returns) {
  # Create portfolio specification
  port_spec <- portfolio.spec(colnames(returns))
  port_spec <- add.constraint(portfolio = port_spec, type = "full_investment")
  port_spec <- add.constraint(portfolio = port_spec, type = "long_only")
  port_spec <- add.objective(portfolio = port_spec, type = "risk", name = "StdDev")
  port_spec <- add.objective(portfolio = port_spec, type = "return", name = "mean")

  # Optimize the portfolio
  opt_port <- optimize.portfolio(R = returns, portfolio = port_spec, optimize_method = "DEoptim", trace = TRUE)

  # Generate efficient frontier
  eff_frontier <- create.EfficientFrontier(R = returns, portfolio = port_spec, n.portfolios = 100, type = "mean-StdDev")

  return(eff_frontier)
}

efficient_frontier1 <- generate_efficient_frontier(portfolio1_combined)
efficient_frontier2 <- generate_efficient_frontier(portfolio2_combined)

# Plot the efficient frontiers
plot(efficient_frontier1, main = "Efficient Frontier - Portfolio 1")
plot(efficient_frontier2, main = "Efficient Frontier - Portfolio 2")
```






# END ~~~~~~~~~~~~~~~~~~~~~

The $opt_10$ object contains the results of the portfolio optimization process performed by the optimize.portfolio function from the PortfolioAnalytics package. It is a complex list that store various details about the optimized portfolio, including the optimized weights, returns, risks, and additional information used and generated during the optimization process.

1. Optimized Weights
The primary component of interest in this object is the optimized weights for each asset in the portfolio. These weights indicate the proportion of the total investment allocated to each asset to achieve the optimization objectives.

2. Expected Returns and Risks
It contains information about the expected returns and risks (standard deviation) of the optimized portfolios. The optimization process aims to find the optimal balance between return and risk based on the specified objectives and constraints.
You sent
3. Constraints and Objectives
The object stores the constraints and objectives defined for the optimization, such as the full investment constraint, box constraints on asset weights, and the objectives for maximizing returns and minimizing risk.

4. Optimization Method and Results
Information about the optimization method used (e.g., "ROI" for R Optimization Infrastructure) and the results of the optimization process, including convergence details and performance metrics.

5. Call and Data
The original function call and the data used for the optimization is also stored in this object. This includes the historical returns data and the portfolio specification.
<<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>>
Parameters:
Portfolio1 = ['acen','allhc','apx','cnpf','bpi','cnvrg','fgen','jgs','meg','pxp']
Portfolio2 = ['agi','ali','bdo','ict','mac','wlcon','nikl','smc','cbc','psei']
Rf = 0.06275 ~ 6.275% (worldgovernmentbonds.com/bond-historical-data/philippines/10-years/) 10 year-PH Riskfree rate
Date range = 01012015-12312023
Count = approx 2000 rows by 20 variables
Data = refinitiv eikon

Use log returns to create a portfolio of the 10 assets:
 a) Optimized Portfolio with No Rebalancing
 b) Optimized Portfolio with Rebalancing
 Create a table comparing the risk and return for each case together with the Sharpe ratio
 You have to use the current risk free rate in the market (10 YEAR TREASURY IS ACCEPTABLE)
 Create the Efficient frontier for both Cases

10 year - Philippines Risk free rate
worldgovernmentbonds.com/bond-historical-data/philippines/10-years/
<<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>>

Guide steps

Collect historical price data for the selected assets.
Calculate log returns for each asset.
Compute the mean returns and covariance matrix.
Define the objective function for portfolio optimization.
Set constraints for the optimization problem.
Implement the quadratic optimization algorithm.
Generate the efficient frontier by varying the target return.
Create optimized portfolios with and without rebalancing.
Calculate risk, return, and Sharpe ratio for each portfolio.
Visualize the efficient frontier for both cases.
Analyze and compare the results.
Prepare a comprehensive report with findings and insights.





















































