---
title: "AL1"
author: "Richard Gan"
date: "`r Sys.Date()`"
output: html_document
---
## Use 10 and 20 portfolio


```{r, initialize}
require(readxl)
require(xts)
require(readr)
require(tidyverse)
require(broom)
require(stargazer)
require(tidyquant)
require(purrr)
require(tidyr)
require(dplyr)
require(quantmod)
require(PerformanceAnalytics)
require(PortfolioAnalytics)
require(TTR)
require(DEoptim)
require(ROI)
#require(qmao)
require(zoo)
require(quadprog)
require(ggplot2)
require(foreach)
#install.packages("ROI.plugin.quadprog")
library(ROI.plugin.quadprog)
#install.packages("ROI.plugin.glpk")
library(ROI.plugin.glpk)
```

# Code starts around line 110

<<<-------------------------------Pull 20 stocks into 2 xlsx (Pre data transformation DONT RUN) ------------------------------->>>
# 1 Data transformation

```{r,eval=FALSE}
# Define portfolios
portfolio1 <- c("acen", "allhc", "apx", "cnpf", "bpi", "cnvrg", "fgen", "jgs", "meg", "pxp")
portfolio2 <- c("agi", "ali", "bdo", "ict", "mac", "wlcon", "nikl", "smc", "cbc", "psei")
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/"
```

```{r,transform1,eval=FALSE}
# 20 downloaded xlsx files acquired from refinitiv eikon
# for loop logic: (1) file name aliased as 'ticker' name (2) drop all fields except exchange date and close and append Ticker name (3) 'ticker' is then mapped to portfolio1 or portfolio2, see list above (4) loop repeats until all tickers are bucketed (5) output is a large list of two portfolios with 10 stocks each
file_list <- list.files(path = path, pattern = "*.xlsx", full.names = TRUE)

df_list_portfolio1 <- list()
df_list_portfolio2 <- list()

for (file in file_list) {
  df <- read_excel(file)
  ticker <- tools::file_path_sans_ext(basename(file)) 
  df <- df %>% select(`Exchange Date`, Close) %>% mutate(Ticker = ticker) 
  
  if (ticker %in% portfolio1) {
    df_list_portfolio1 <- append(df_list_portfolio1, list(df))
  } else if (ticker %in% portfolio2) {
    df_list_portfolio2 <- append(df_list_portfolio2, list(df))
  }
}
df_list_portfolio1
df_list_portfolio2
```

```{r,transform2,eval=FALSE}
# merge the two large list into a two dataframes
port1 <- data.frame(bind_rows(df_list_portfolio1))
port2 <- data.frame(bind_rows(df_list_portfolio2))
port1
port2
```

```{r,transform3, eval=FALSE}
# merge again two form 11 fields with close as the price and ticker as the column
port1 <- port1 %>%
  select(Exchange.Date, Ticker, Close) %>%
  pivot_wider(names_from = Ticker, values_from = Close)
port2 <- port2 %>%
  select(Exchange.Date, Ticker, Close) %>%
  pivot_wider(names_from = Ticker, values_from = Close)

# rename
colnames(port1)[1] <- "date"
colnames(port2)[1] <- "date"

# write to output file for data submission, this file is used below. it is saved in your setwd or working directory
#write.csv(port1, "port1.csv",row.names = FALSE) 
#write.csv(port2, "port2.csv",row.names = FALSE) 


port2
port1
```

# DUE TO ISSUES WITH CSV FORMAT PORT1 AND PORT2 CSV HAS BEEN SAVED INTO XLSX
# The two files (port1.xlsx,port2.xlsx)is subsequently loaded in the below code. PLEASE USE BELOW CHUNKS AS YOUR STARTING POINT

<<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>>

<<<-------------------------------Start here for Project------------------------------->>>
# 2 Import Data

```{r, Start Here}
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port1.xlsx"
port1 <- read_xlsx(path)

path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port2.xlsx"
port2 <- read_xlsx(path)
```
<<<-------------------------------Data Transformation PORTION------------------------------->>>

```{r}
# Convert into df
port1 <- data.frame(port1)
port2 <- data.frame(port2)
head(port1)

```
```{r}
# Check data type
str(port1)
str(port2)
```

```{r}
# Check missing
sum(is.na(port1))
sum(is.na(port2))
```

```{r}
# Convert to numeric except date
port1[, -1] <- lapply(port1[, -1], function(x) as.numeric(as.character(x)))
port2[, -1] <- lapply(port2[, -1], function(x) as.numeric(as.character(x)))
```
```{r}
sum(is.na(port1))
sum(is.na(port2))
```

```{r}
# Drop missing
port1 <- na.omit(port1)
port2 <- na.omit(port2)
#port1 <- port1[complete.cases(port1), ]
#port2 <- port2[complete.cases(port2), ]
sum(is.na(port1))
sum(is.na(port2))
```

```{r}
port1 <- xts(port1[,-1], order.by = as.Date(port1$date))
port2 <- xts(port2[,-1], order.by = as.Date(port2$date))
head(port1)
```

```{r}
# Check data type
str(port1)
str(port2)
```
```{r}
sum(is.na(port1))
sum(is.na(port2))
```

```{r}
head(port1)
head(port2)
```

<<<-------------------------------RETURNS PORTION------------------------------->>>

$$
r_p \sum w_i~r_i
$$

<<<-------------------------------Log returns------------------------------->>>
```{r,log return}
logReturns1 <- PerformanceAnalytics::Return.calculate(port1,method="log")
logReturns2 <- PerformanceAnalytics::Return.calculate(port2,method="log")

# we get the log and get the difference which is equivalent to getting the log first and difference
```

```{r}
logReturns1 <- logReturns1[-1]
logReturns2 <- logReturns2[-1]
```

<<<-------------------------------Log returns w na.omit------------------------------->>>

```{r,log return,eval=FALSE}
logReturns1na <- na.omit(CalculateReturns(port1, method = "log"))
logReturns2na <- na.omit(CalculateReturns(port2, method = "log"))
view(logReturns2na)
```

<<<-------------------------------Log returns for single stocks ------------------------------->>>
```{r, log return v2 - Only for single returns,eval=FALSE}
logreturns_single1 <- periodReturn(port1,
             period="daily"
             ,subset = NULL
             ,type = "log"
             ,leading = TRUE)
logreturns_single2 <- periodReturn(port2,
             period="daily"
             ,subset = NULL
             ,type = "log"
             ,leading = TRUE)
view(logreturns_single1)
```
```{r,eval=FALSE}
# Calculate daily returns
returns <- ROC(port1, type = "continuous")[-1]
```


<<<-------------------------------Log returns for two stocks ------------------------------->>>


```{r,port1 for 2 stocks w 50/50 ratio,eval=FALSE}
logreturns_double1 <- na.omit(port1yr)
logreturns_double1 <- logreturns_double1[,1:2]
```

```{r}
logreturns_double2 <- na.omit(port2yr)
logreturns_double2 <- logreturns_double2[,1:2]
```


```{r,port1 for 2 stocks w 50/50 ratio}
port1ret <- PerformanceAnalytics::Return.portfolio(logreturns_double1,c(0.5,0.5))
view(port1ret)
colnames(port1ret) <- "port1_50_Ret"
mean(port1ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```

```{r,port2 for 2 stocks w 50/50 ratio}

port2ret <- PerformanceAnalytics::Return.portfolio(logreturns_double2,c(0.5,0.5))
view(port2ret)
colnames(port2ret) <- "port2_50_Ret"
mean(port2ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```


<<<-------------------------------Handling Infinity------------------------------->>>

```{r, Inf Checker function}
# Code is breaking due to instances of infinity when a 0 is returned and is subsequently divided by a function

# initialize infinity check function
infChecker <- function(data) {
  infIndices <- which(is.infinite(data), arr.ind = TRUE)
  return(infIndices)
}

# var list
dataList <<- list(logreturns_single1
           ,logreturns_single2
           ,logReturns1
           ,logReturns2
           ,logReturns1na
           ,logReturns2na
           ,port1
           ,port2
           ,logreturns_double1
           ,logreturns_double2
           ,port1ret
           ,port2ret
           ,port1yr
           ,port2yr
           ) 
vrList <<- c("logreturns_single1"
           ,"logreturns_single2"
           ,"logReturns1"
           ,"logReturns2"
           ,"logReturns1na"
           ,"logReturns2na"
           ,"port1"
           ,"port2"
           ,"logreturns_double1"
           ,"logreturns_double2"
           ,"port1ret"
           ,"port2ret"
           ,"port1yr"
           ,"port2yr"
           ) 

# Apply f(x)
infIndices <- lapply(dataList, infChecker)
names(infIndices) <- vrList
#print(infIndices)
```


```{r, handling Infinity}
# Here from the above code we can see both port1ret and port2ret variable is causing the code to break
# Inf value is breaking code due to division by 0
#port2yrInf <- which(is.infinite(port2ret), arr.ind = TRUE)

lapply(port2ret, infChecker)
lapply(port1ret, infChecker)

```


```{r}
port1ret[is.infinite(port1ret)] <- NA
port1ret <- port1ret[complete.cases(port1ret), ]
lapply(port1ret, infChecker)
```

```{r, handling Infinity}
# Use NA and Drop
port2ret[is.infinite(port2ret)] <- NA
port2ret <- port2ret[complete.cases(port2ret), ]
lapply(port2ret, infChecker)
```


<<<-------------------------------Rebal Continued------------------------------->>>

```{r,port1 for 2 stocks w 50/50 ratio}
#logreturns_double1 <- na.omit(port1yr)
#logreturns_double1 <- logreturns_double1[,1:2]
#port1ret <- PerformanceAnalytics::Return.portfolio(logreturns_double1,c(0.5,0.5))
view(port1ret)
colnames(port1ret) <- "port1_50_Ret"
mean(port1ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```

```{r,port2 for 2 stocks w 50/50 ratio}
# Re initialize without the Inf values
#logreturns_double2 <- na.omit(port2yr)
#logreturns_double2 <- logreturns_double2[,1:2]
#port2ret <- PerformanceAnalytics::Return.portfolio(logreturns_double2,c(0.5,0.5))
view(port2ret)
colnames(port2ret) <- "port2_50_Ret"
mean(port2ret)*1512 # 2016 == 8 years, 252 == 1  year trading days
```


```{r,rebal_port1_doublestocks}
port1_50 <- na.omit(port1yr)
port1_50 <- port1_50[,1:2]
port1RetRebal <- Return.portfolio(port1_50, c(0.5,0.5),rebalance_on = "months")
view(port1RetRebal)
colnames(port1RetRebal) <- "port50Rebal_Ret"
```

```{r}
head(port1ret,30)
head(port1RetRebal,30)
```

```{r,rebal_port2_doublestocks}
port2_50 <- na.omit(port2yr)
port2_50 <- port1_50[,1:2]
port2RetRebal <- Return.portfolio(port2_50, c(0.5,0.5),rebalance_on = "months")
view(port2RetRebal)
colnames(port2RetRebal) <- "port50Rebal_Ret"
```

```{r}
head(port2ret,30)
head(port2RetRebal,30)
```


```{r}
allport1ret <- cbind(port1ret,port1RetRebal)
colnames(allport1ret) <- c("Non-Rebalanced","Monthly Rebalanced")
table.AnnualizedReturns(allport1ret,Rf=0.1/2016)
```






# 3 Exploratory Data Analysis (EDA)

## 3.1 EDA Portfolio1
### 3.1.1 Descriptive Statistics - mean, std dev, skewness and kurtosis of asset returns
## 3.1.1 Correlation matrix of asset returns.
## 3.1.3Visualization - time series plots, histograms, density plots of asset returns; HEATMAP of the correlation matrix

```{r,EDA Portfolio1}

```


## 3.2 EDA Portfolio2
### 3.2.1 Descriptive Statistics - mean, std dev, skewness and kurtosis of asset returns
### 3.2.2 Correlation matrix of asset returns.
### 3.2.3 Visualization - time series plots, histograms, density plots of asset returns; HEATMAP of the correlation matrix

```{r,EDA Portfolio2}

```


# Optimization, Rebalancing, Effient Frontier


# 4 Modeling
## 4.1 Portfolio 1
### 4.1.1 Optimized Portfolio with No Rebalancing
### 4.1.2 Optimized Portfolio with Rebalancing








#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~


Port optimization

```{r, Start Here,eval=FALSE}
path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port1.xlsx"
port1 <- read_xlsx(path)

path <- "C:/Users/chard/OneDrive/Desktop/MSCF/Courses/Term5/Trading Software and Programming/scripts/PortOpt_Final/port2.xlsx"
port2 <- read_xlsx(path)
```
<<<-------------------------------Data Transformation PORTION------------------------------->>>

```{r}
# Convert into df
port1 <- data.frame(port1)
port2 <- data.frame(port2)
```

```{r}
# Convert to numeric except date
port1[, -1] <- lapply(port1[, -1], function(x) as.numeric(as.character(x)))
port2[, -1] <- lapply(port2[, -1], function(x) as.numeric(as.character(x)))
```


```{r}
# Drop missing
port1 <- na.omit(port1)
port2 <- na.omit(port2)
#port1 <- port1[complete.cases(port1), ]
#port2 <- port2[complete.cases(port2), ]
sum(is.na(port1))
sum(is.na(port2))
```

```{r}
port1 <- xts(port1[,-1], order.by = as.Date(port1$date))
port2 <- xts(port2[,-1], order.by = as.Date(port2$date))
head(port1)
```

```{r}
stockReturns <- PerformanceAnalytics::Return.calculate(port1,method="log")
port1 <- stockReturns[-1]
```


```{r}
port_spec1 <- portfolio.spec(assets = colnames(port1))
port_spec1
View(port_spec1)
```

# The portspec object holds the Constraints
<<--------------------------------->>

```{r}
#min=0.99 max1.01 to relax constraint

port_spec1 <- add.constraint(portfolio=port_spec1,type="weight_sum"
                             ,min_sum=0.0
                             ,max_sum=1.0)

```

```{r}
port_spec1 <- add.constraint(portfolio = port_spec1
                             ,type = "long_only")
```

```{r}
port_spec1 <- add.constraint(portfolio = port_spec1
                             ,type = "box"
                             , min=0.0
                             , max=1.0)
```

# the portMinVar1 holds the Objectives which in this case is minimize risk
<<--------------------------------->>


```{r}
# Can be minimize risk or max return
# For max specify type = "return" or type = "risk"
portMinVar1 <- port_spec1
portMinVar1 <- add.objective(portfolio = portMinVar1
                             ,type= "risk"
                             ,name = "StdDev")
View(portMinVar1)
```

```{r}
minVarOpt1 <- optimize.portfolio(R = port1
                                 , portfolio = portMinVar1
                                 , optimize_method = "random"
                                 )
```

```{r}
(minVarOpt1_Weights <- extractWeights(minVarOpt1))
```

```{r}
minVarOpt1_Return <- Return.portfolio(R = port1, 
                                weights = minVarOpt1_Weights)
```



# Quadtratic Programming

```{r}
quadProg1 <- optimize.portfolio(R = port1
                                 , portfolio = portMinVar1
                                 , optimize_method = "quadprog"
                                 ,verbose=TRUE
                                 )
```

```{r}
(quadProg1_Weights <- extractWeights(minVarOpt1))
```


#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~







<<<-------------------------------Portfolio specifications / Constraints------------------------------->>>

```{r}
# Create initial portfolio specification
portfolio_spec <- portfolio.spec(assets = port1)

# Add constraint for full investment
portfolio_spec <- add.constraint(portfolio = portfolio_spec, 
                                 type = "full_investment")

# Add constraint for long only positions
portfolio_spec <- add.constraint(portfolio = portfolio_spec, 
                                 type = "long_only")
# Add weight constraints (e.g., between 5% and 40% for each asset)
portfolio_spec <- add.constraint(portfolio = portfolio_spec, type = "box", min = 0.0, max = 1.0)

#Use box in your output as per recommendation
#Long only minimum 0 and .05
# 0 is unnecessary if there is 0.5
```

<<<-------------------------------Objective Function------------------------------->>>

```{r}

# Objective to maximize return
portfolio_max_return <- add.objective(portfolio = portfolio_spec, 
                                      type = "return", name = "mean")
summary(portfolio_max_return)
```

```{r}
# Objective to minimize risk (standard deviation)
portfolio_min_risk <- add.objective(portfolio = portfolio_spec, 
                                    type = "risk", name = "StdDev")
```

```{r}
# Objective for combination of risk and return
portfolio_risk_return <- add.objective(portfolio = portfolio_spec, 
                                       type = "return", name = "mean")
portfolio_risk_return <- add.objective(portfolio = portfolio_risk_return,
                                       type = "risk", name = "StdDev")
```

```{r}
#0.5 might not be possible of 100 is 20 stocks
#contraint is ==0.1 to 0.3
#What weare interested in is the profile of minimization and maximization for us to create the efficient frontier
```


```{r}
# Optimize portfolio for combination of risk and return
opt_risk_return <- optimize.portfolio(R = returns, portfolio = portfolio_risk_return, optimize_method = "ROI", trace = TRUE)
```

<<<-------------------------------Weights------------------------------->>>

```{r}
# Extract weights
(weights_max_return <- extractWeights(opt_max_return))
(weights_min_risk <- extractWeights(opt_min_risk))
(weights_risk_return <- extractWeights(opt_risk_return))
```


```{r}
# CalculateReturns()
# you will get weights on each assets when calculating returns, 1000 rows x 5 will be multiplied 
# you will get portfolio.returns as weighted returns
# each row is a weighted sum of the W1X1 + W2X2 where W == weights, X== asset
```

```{r}
# Calculate portfolio returns
portfolio_returns_max <- Return.portfolio(R = returns, 
                                weights = weights_max_return)

portfolio_returns_min <- Return.portfolio(R = returns, 
                                weights = weights_min_risk)

portfolio_returns_risk_return <- Return.portfolio(R = returns,
                                weights = weights_risk_return)
```

```{r}
# Plot cumulative returns
chart.CumReturns(portfolio_returns_max, main = "Cumulative Returns - Max Return Portfolio", wealth.index = TRUE)
chart.CumReturns(portfolio_returns_min, main = "Cumulative Returns - Min Risk Portfolio", wealth.index = TRUE)
chart.CumReturns(portfolio_returns_risk_return, main = "Cumulative Returns - Risk/Return Portfolio", wealth.index = TRUE)
```



```{r}
# Calculate portfolio returns
portfolio_returns_max <- Return.portfolio(R = returns, 
                                weights = weights_max_return)

portfolio_returns_min <- Return.portfolio(R = returns, 
                                weights = weights_min_risk)

portfolio_returns_risk_return <- Return.portfolio(R = returns,
                                weights = weights_risk_return)
```

```{r}
# Plot cumulative returns
chart.CumReturns(portfolio_returns_max, main = "Cumulative Returns - Max Return Portfolio", wealth.index = TRUE)
chart.CumReturns(portfolio_returns_min, main = "Cumulative Returns - Min Risk Portfolio", wealth.index = TRUE)
chart.CumReturns(portfolio_returns_risk_return, main = "Cumulative Returns - Risk/Return Portfolio", wealth.index = TRUE)
```




## 4.2 Portfolio 2
### 4.2.1 Optimized Portfolio with No Rebalancing
### 4.2.2 Optimized Portfolio with Rebalancing

```{r}
break
Repeat above code
```



# 5. Comparison of the Optimized Portfolios and Discussion




# 6. Efficient Frontier for both Cases

```{r}
?create.EfficientFrontier
```

```{r}
#Generate Efficient Frontier

# Function to generate multiple points on the efficient frontier
generate_efficient_frontier <- function(returns, portfolio_spec, num_points = 50) {
  eff_frontier <- data.frame(Return = numeric(), Risk = numeric(), SharpeRatio = numeric())
  
  min_return <- min(colMeans(returns))
  max_return <- max(colMeans(returns))
  
  target_returns <- seq(min_return, max_return, length.out = num_points)
  
  for (target_return in target_returns) {
    tmp_portfolio <- portfolio_spec
    
    # Add the return objective dynamically for each target return
    tmp_portfolio <- add.constraint(tmp_portfolio, type = "return", return_target = target_return)
    tmp_portfolio <- add.objective(tmp_portfolio, type = "risk", name = "StdDev")
    
    opt <- optimize.portfolio(R = returns, portfolio = tmp_portfolio, optimize_method = "ROI", trace = TRUE)
    
    ret <- mean(Return.portfolio(R = returns, weights = extractWeights(opt)))
    risk <- StdDev(Return.portfolio(R = returns, weights = extractWeights(opt)))
    sharpe <- ret / risk
    
    eff_frontier <- rbind(eff_frontier, data.frame(Return = ret, Risk = risk, SharpeRatio = sharpe))
  }
  
  return(eff_frontier)
}

```

```{r}
# Generate efficient frontier
efficient_frontier <- generate_efficient_frontier(returns, portfolio_spec, num_points = 100)

# Plot efficient frontier
plot(efficient_frontier$Risk, efficient_frontier$Return, type = "l", col = "blue", lwd = 2,
     xlab = "Risk (StdDev)", ylab = "Return", main = "Efficient Frontier")
points(efficient_frontier$Risk, efficient_frontier$Return, col = "blue", pch = 16)

```


<<<-------------------------------model1 EF------------------------------->>>


```{r, Efficient Frontier Model 1}

# Calculate Efficient Frontier
calculate_efficient_frontier <- function(returns) {
  returns_matrix <- returns %>% select(-date) %>% as.matrix()
  colnames(returns_matrix) <- c("port1", "port2")
  mean_returns <- colMeans(returns_matrix)
  cov_matrix <- cov(returns_matrix)
  
  eff_frontier <- portfolio.optim(returns = mean_returns, covmat = cov_matrix, shorts = FALSE)
  return(eff_frontier)
}

efficient_frontier <- calculate_efficient_frontier(eff_frontier)

# Plot Efficient Frontier
plot_efficient_frontier <- function(eff_frontier) {
  eff_df <- data.frame(
    Risk = sqrt(diag(eff_frontier$covmat)),
    Return = eff_frontier$pm
  )
  
  ggplot(eff_df, aes(x = Risk, y = Return)) +
    geom_point(color = "blue", size = 3) +
    geom_line(color = "blue") +
    labs(title = "Efficient Frontier",
         x = "Risk (Standard Deviation)",
         y = "Return") +
    theme_minimal()
}

plot_efficient_frontier(efficient_frontier)

```


<<<-------------------------------model2 EF------------------------------->>>

```{r}
#Generate Efficient Frontier



# Function to generate multiple points on the efficient frontier
generate_efficient_frontier <- function(returns, portfolio_spec, num_points = 50) {
  eff_frontier <- data.frame(Return = numeric(), Risk = numeric(), SharpeRatio = numeric())
  
  min_return <- min(colMeans(returns))
  max_return <- max(colMeans(returns))
  
  target_returns <- seq(min_return, max_return, length.out = num_points)
  
  for (target_return in target_returns) {
    tmp_portfolio <- portfolio_spec
    
    # Add the return objective dynamically for each target return
    tmp_portfolio <- add.constraint(tmp_portfolio, type = "return", return_target = target_return)
    tmp_portfolio <- add.objective(tmp_portfolio, type = "risk", name = "StdDev")
    
    opt <- optimize.portfolio(R = returns, portfolio = tmp_portfolio, optimize_method = "ROI", trace = TRUE)
    
    ret <- mean(Return.portfolio(R = returns, weights = extractWeights(opt)))
    risk <- StdDev(Return.portfolio(R = returns, weights = extractWeights(opt)))
    sharpe <- ret / risk
    
    eff_frontier <- rbind(eff_frontier, data.frame(Return = ret, Risk = risk, SharpeRatio = sharpe))
  }
  
  return(eff_frontier)
}

```

```{r}
# Generate efficient frontier
efficient_frontier <- generate_efficient_frontier(returns, portfolio_spec, num_points = 100)

# Plot efficient frontier
plot(efficient_frontier$Risk, efficient_frontier$Return, type = "l", col = "blue", lwd = 2,
     xlab = "Risk (StdDev)", ylab = "Return", main = "Efficient Frontier")
points(efficient_frontier$Risk, efficient_frontier$Return, col = "blue", pch = 16)

```

<<<-------------------------------model3 EF------------------------------->>>
```{r}
# Function to extract and calculate returns
calculate_returns <- function(data) {
  if (is.null(data) || nrow(data) == 0 || 
      !"Exchange Date" %in% names(data) || !"Close" %in% names(data)) {
    warning("Invalid data structure")
    return(NULL)
  }
  
  data %>%
    select(`Exchange Date`, Close) %>%
    arrange(`Exchange Date`) %>%
    mutate(Returns = (Close - lag(Close)) / lag(Close)) %>%
    select(`Exchange Date`, Returns) %>%
    na.omit()
}

# Process both portfolios
process_portfolio <- function(portfolio_stocks) {
  portfolio_data <- map(stock_data[portfolio_stocks], calculate_returns)
  portfolio_data <- compact(portfolio_data)  # Remove any NULL results
  
  # Combine all stocks' returns into a single dataframe
  reduce(portfolio_data, function(acc, x) {
    full_join(acc, x, by = "Exchange Date") %>%
      rename(!!sym(names(which(portfolio_stocks == cur_data()$name))) := Returns)
  })
}

# Calculate returns for both portfolios
portfolio1_combined <- process_portfolio(portfolio1)
portfolio2_combined <- process_portfolio(portfolio2)

# Check the results
print(dim(portfolio1_combined))
print(names(portfolio1_combined))
print(dim(portfolio2_combined))
print(names(portfolio2_combined))
```


```{r}
library(dplyr)
library(purrr)
library(tidyr)

# Function to extract and calculate returns
calculate_returns <- function(data) {
  if (is.null(data) || nrow(data) == 0 ||
      !"Exchange Date" %in% names(data) || !"Close" %in% names(data)) {
    warning("Invalid data structure")
    return(NULL)
  }
  
  data %>%
    select(`Exchange Date`, Close) %>%
    arrange(`Exchange Date`) %>%
    mutate(Returns = (Close - lag(Close)) / lag(Close)) %>%
    select(`Exchange Date`, Returns) %>%
    na.omit()
}

# Process both portfolios
process_portfolio <- function(portfolio_stocks, stock_data) {
  portfolio_data <- map(stock_data[portfolio_stocks], calculate_returns)
  portfolio_data <- compact(portfolio_data) # Remove any NULL results
  
  # Combine all stocks' returns into a single dataframe
  combined_data <- reduce(portfolio_data, function(acc, x) {
    full_join(acc, x, by = "Exchange Date")
  })
  
  combined_data
}

# Example stock data (replace with your actual data)
stock_data <- list(
  stock1 = data.frame(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 100, 200)),
  stock2 = data.frame(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 50, 150)),
  stock3 = data.frame(`Exchange Date` = as.Date('2022-01-01') + 0:9, Close = runif(10, 200, 300))
)

portfolio1 <- c("stock1", "stock2")
portfolio2 <- c("stock2", "stock3")

# Calculate returns for both portfolios
portfolio1_combined <- process_portfolio(portfolio1, stock_data)
portfolio2_combined <- process_portfolio(portfolio2, stock_data)

# Check the results
print(dim(portfolio1_combined))
print(names(portfolio1_combined))
print(dim(portfolio2_combined))
print(names(portfolio2_combined))

```


```{r}
# Step 5: Generate the efficient frontier
generate_efficient_frontier <- function(returns) {
  # Create portfolio specification
  port_spec <- portfolio.spec(colnames(returns))
  port_spec <- add.constraint(portfolio = port_spec, type = "full_investment")
  port_spec <- add.constraint(portfolio = port_spec, type = "long_only")
  port_spec <- add.objective(portfolio = port_spec, type = "risk", name = "StdDev")
  port_spec <- add.objective(portfolio = port_spec, type = "return", name = "mean")

  # Optimize the portfolio
  opt_port <- optimize.portfolio(R = returns, portfolio = port_spec, optimize_method = "DEoptim", trace = TRUE)

  # Generate efficient frontier
  eff_frontier <- create.EfficientFrontier(R = returns, portfolio = port_spec, n.portfolios = 100, type = "mean-StdDev")

  return(eff_frontier)
}

efficient_frontier1 <- generate_efficient_frontier(portfolio1_combined)
efficient_frontier2 <- generate_efficient_frontier(portfolio2_combined)

# Plot the efficient frontiers
plot(efficient_frontier1, main = "Efficient Frontier - Portfolio 1")
plot(efficient_frontier2, main = "Efficient Frontier - Portfolio 2")
```


# Appendix
<<<-------------------------------END------------------------------->>>

<<<-------------------------------GUIDE------------------------------->>>


The $opt_10$ object contains the results of the portfolio optimization process performed by the optimize.portfolio function from the PortfolioAnalytics package. It is a complex list that store various details about the optimized portfolio, including the optimized weights, returns, risks, and additional information used and generated during the optimization process.

1. Optimized Weights
The primary component of interest in this object is the optimized weights for each asset in the portfolio. These weights indicate the proportion of the total investment allocated to each asset to achieve the optimization objectives.

2. Expected Returns and Risks
It contains information about the expected returns and risks (standard deviation) of the optimized portfolios. The optimization process aims to find the optimal balance between return and risk based on the specified objectives and constraints.
You sent
3. Constraints and Objectives
The object stores the constraints and objectives defined for the optimization, such as the full investment constraint, box constraints on asset weights, and the objectives for maximizing returns and minimizing risk.

4. Optimization Method and Results
Information about the optimization method used (e.g., "ROI" for R Optimization Infrastructure) and the results of the optimization process, including convergence details and performance metrics.

5. Call and Data
The original function call and the data used for the optimization is also stored in this object. This includes the historical returns data and the portfolio specification.
<<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>>
Parameters:
Portfolio1 = ['acen','allhc','apx','cnpf','bpi','cnvrg','fgen','jgs','meg','pxp']
Portfolio2 = ['agi','ali','bdo','ict','mac','wlcon','nikl','smc','cbc','psei']
Rf = 0.06275 ~ 6.275% (worldgovernmentbonds.com/bond-historical-data/philippines/10-years/) 10 year-PH Riskfree rate
Date range = 01012015-12312023
Count = approx 2000 rows by 20 variables
Data = refinitiv eikon

Use log returns to create a portfolio of the 10 assets:
 a) Optimized Portfolio with No Rebalancing
 b) Optimized Portfolio with Rebalancing
 Create a table comparing the risk and return for each case together with the Sharpe ratio
 You have to use the current risk free rate in the market (10 YEAR TREASURY IS ACCEPTABLE)
 Create the Efficient frontier for both Cases

10 year - Philippines Risk free rate
worldgovernmentbonds.com/bond-historical-data/philippines/10-years/
<<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>><<<<<<<<<<<<>>>>>>>>>>

Guide steps

Collect historical price data for the selected assets.
Calculate log returns for each asset.
Compute the mean returns and covariance matrix.
Define the objective function for portfolio optimization.
Set constraints for the optimization problem.
Implement the quadratic optimization algorithm.
Generate the efficient frontier by varying the target return.
Create optimized portfolios with and without rebalancing.
Calculate risk, return, and Sharpe ratio for each portfolio.
Visualize the efficient frontier for both cases.
Analyze and compare the results.
Prepare a comprehensive report with findings and insights.





























