---
title: "technicalAnalysis"
author: "Richard Gan"
date: "`r Sys.Date()`"
output: html_document
---



$$
\begin{aligned}&y_i = \beta_0 + \beta_2 x_{i2} + \beta_3 xx{i3} + ... + \beta _n x_{in} + \epsilon_i

\\&\textbf{where, for } i = n~\textbf{observations:}
\\&y_i\text{response variable}
\\&x_i\text{explanatory variables}
\\&\beta_0=\text{y-intercept (constant)}
\\&\beta_n=\text{slope coefficients for each explanatory variable}
\\&\epsilon_i\text{the model's error term (aka residuals)}
\end{aligned}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(quantmod, TTR, PerformanceAnaytics,PortfolioAnalytics)

```

```{r}

require(pacman)
```
```{r}
#hw 1 each on output, use 5 years  | trend, volatility, and price
# show the formula using latex for your measure
#do not use more than one per measure (vol, price)
```


```{r}
getSymbols('AAPL',src='yahoo',from='2023-01-01',to='2023-12-31'
)
apple_cl<- cl(APPL); view(apple_cl)
```


First measure
Simple Moving Average (SMA) - this is calculated by averaging the prices over a specific number of periods.

Advantages: Easy to calculate, smooths out price data.
Disadvantages: Lags behind current price trends.
Use in Trading: Identifies trends. When the price is above the SMA, it's an uptrend; below, it's a downtrend.

$$
\text{SMA}_n = \frac{1}{n} \sum_{i=1}^{n} P_i
$$
$$
\begin{aligned}
\text{where}\
  & P_i && \text{: price at time i} \\
  & \text{n} && \text{: number of periods} \\
  \end{aligned}
$$



```{r}
mySMA <- function (price,n){
  sma <- c()
  sma[1:(n-1)] <- NA
  for (i in n:length(price)){
    sma[i]<-mean(price[(i-n+1):i])
  }
  sma <- reclass(sma,price)
  return(sma)
}
```

apply function
```{r}
mysma <- mySMA(aapl_cl,n=20)
  head(mysma,n=22); tail(mysma,n=5)
```




apply function
```{r}
sma <- TTR:SMA(aapl_cl,n=20)
head(mysma,n=22); tail(mysma,n=5)
  head(mysma,n=22); tail(mysma,n=5)
```



Exponential Moving Average (EMA) - this is similar to SMA but gives more weight to recent prices.

Advantages: More responsive to recent price changes than SMA.
Disadvantages: Can be more volatile than SMA.
Use in Trading: Identifies trends with less lag. Often used in conjunction with other indicators.

$$
\text{EMA}_t = \alpha \cdot P_t + (1 - \alpha) \cdot \text{EMA}_{t-1}
$$

$$
\begin{aligned}
\text{where}\
  & P_t && \text{: price at time t} \\
  & \alpha && \text{: Smoothing operator = 2/(ð‘›+ 1)} \\
  & \text{n} && \text{: number of periods} \\
  \end{aligned}
$$

```{r}
ema <-  
```
```{r}
chartSeries(AAPL,TA=NULL,theme=ChartTheme("white"),name="Apple Inc.")
addTA(sma,on=1,col='red')
addTA(ema,on=1,col='red')
```

 

```{r}
ema_20 <- EMA(prices,n=20)
sma_50 <- EMA(prices,n=20)
```

```{r}

chartSeries(AAPL,TA=NULL,theme=charTheme("white"),name='Apple  Inc(AAPL)'
            
```
## SMA Plot
It is the average price over the last n periods.
It smooths out price data by creating a constantly updated average price.
It helps identify the direction of the trend.

### What May be gleaned from the SMA plot:

* When the price is above the SMA, it indicates a potential uptrend.
* When the price is below the SMA, it indicates a potential downtrend.
* Crossovers (when the price crosses the SMA) can signal potential buy or sell opportunities.

## EMA Plot
The EMA plot gives more weight to recent prices, making it more responsive to new information. It smooths out price fluctuations and highlights trends more quickly than SMA.
### What may be gleaned from the EMA plot:
Similar to SMA, but more sensitive to recent price changes.
Crossovers between EMA and price can signal potential buy or sell opportunities.
Useful for detecting trend reversals sooner than SMA.

Both SMA and EMA help identify the trend direction. When the price is above the moving average, it indicates an uptrend; when below, a downtrend.

Crossover Strategy:

Buy Signal: When the shorter-term moving average (e.g., 20-day EMA) crosses above the longer-term moving average (e.g., 50-day SMA), it indicates a potential uptrend, and you might consider buying. (Buy signal arises when a short-run SMA crosses from below to above a long-run SMA.)

Sell Signal: When the shorter-term moving average crosses below the longer-term moving average, it indicates a potential downtrend, and you might consider selling. (Sell signal arrises when a short-run SMA crosses from above to above a long-run SMA.)
```{r}
#plotting
{r}
chartSeries(AAPL, TA = NULL, theme = chartTheme("white"), name = "Apple Inc. (AAPL)")
addTA(sma, on = 1, col="red")
addTA(ema, on = 1, col="green")
```


```{r}
3) Bollinger Bands - consists of a middle band (SMA) and two outer bands that are standard deviations away from the middle band (usually 2).

Advantages: Indicates volatility, overbought, and oversold conditions.
Disadvantages: Can give false signals in a trending market.
Use in Trading: Buy when prices touch the lower band and sell when they touch the upper band.
$$
\text{Upper Band} = \text{SMA}_n + 2 \cdot \sigma\\
\text{Lower Band} = \text{SMA}_n - 2 \cdot \sigma
$$
$$
\begin{aligned}
\text{where}\
  & SMA_n && \text{: Simple Moving Average over ð‘› periods} \\
  & \sigma && \text{: Simple Moving Average over ð‘› periods}
  \end{aligned}
$$
```
```{r}
4) Relative Strength Index (RSI) - this measures the speed and change of price movements.

Advantages: Identifies overbought and oversold conditions.
Disadvantages: Can stay in overbought/oversold conditions for a long time during strong trends.
Use in Trading: Buy when RSI is below 30 and sell when RSI is above 70.

$$
\text{RSI} = 100 - \left( \frac{100}{1 + \frac{\text{Average Gain}}{\text{Average Loss}}} \right)
$$
$$
\begin{aligned}
\text{where}\
  & \text{Average gain} && \text{: average gain over the specific period}\\
  & \text{Average loss} && \text{: average loss over the specific period}\\
  \end{aligned}
$$
```

